---
title: "Heming HW6"
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Load the libraries
```{r}
library(tidyverse)
library(skimr)
library("RColorBrewer")
```
#Gapminder
```{r}
library("gapminder")
```
Load the data into a new object called gapminder.
```{r}
gapminder <- 
  gapminder::gapminder
```
1. Explore the data using the various function you have learned. Is it tidy, are there any NA's, what are its dimensions, what are the column names, etc.

```{r}
summary(gapminder)
```
```{r}
names(gapminder)
```

```{r}
gapminder %>% 
  summarize(number_nas1=sum(is.na(country)),
            number_nas2=sum(is.na(continent)),
            number_nas3=sum(is.na(year)),
            number_nas4=sum(is.na(lifeExp)),
            number_nas5=sum(is.na(pop)),
            number_nas6=sum(is.na(gdpPercap)))
```
2. We are interested in the relationship between per capita GDP and life expectancy; i.e. does having more money help you live longer on average. Make a quick plot below to visualize this relationship.
```{r}
gapminder %>% 
  ggplot(aes(x=gdpPercap,y=lifeExp))+
  geom_jitter()+
  labs(title = "per capita GDP vs. life expectancy",
       x= "per capita GDP",
       y= "life expectancy")
```

3. There is extreme disparity in per capita GDP. Rescale the x axis to make this easier to interpret. How would you characterize the relationship?
```{r}
gapminder %>% 
  ggplot(aes(x=gdpPercap,y=lifeExp))+
  geom_jitter()+
  scale_x_log10()+
  labs(title = "per capita GDP vs. life expectancy",
       x= "per capita GDP",
       y= "life expectancy")
```
4. This should look pretty dense to you with significant overplotting. Try using a faceting approach to break this relationship down by year.
```{r}
gapminder %>% 
  ggplot(aes(x=gdpPercap,y=lifeExp))+
  geom_jitter()+
  scale_x_log10()+
  labs(title = "per capita GDP vs. life expectancy",
       x= "per capita GDP",
       y= "life expectancy")+
  facet_wrap(~year)
```
5. Simplify the comparison by comparing only 1952 and 2007. Can you come to any conclusions?
```{r}
gapminder %>% 
  filter(year=="1952"|year=="2007") %>%  
  ggplot(aes(x=gdpPercap,y=lifeExp))+
  geom_jitter()+
  scale_x_log10()+
  labs(title = "per capita GDP vs. life expectancy",
       x= "per capita GDP",
       y= "life expectancy")+
  facet_grid(~year)
```
6. Let's stick with the 1952 and 2007 comparison but make some aesthetic adjustments. First try to color by continent and adjust the size of the points by population. Add + scale_size(range = c(0.1, 10), guide = "none") as a layer to clean things up a bit.
```{r}
gapminder %>% 
  filter(year=="1952"|year=="2007") %>%  
  ggplot(aes(x=gdpPercap,y=lifeExp, color=continent, size=pop))+
  geom_jitter()+
  scale_x_log10()+
  scale_size(range = c(0.1, 10), guide = "none")+
  labs(title = "per capita GDP vs. life expectancy",
       x= "per capita GDP",
       y= "life expectancy")+
  facet_grid(~year)
```
7. Although we did not introduce them in lab, ggplot has a number of built-in themes that make things easier. I like the light theme for these data, but you can see lots of options. Apply one of these to your plot above.
```{r}
gapminder %>% 
  filter(year=="1952"|year=="2007") %>%  
  ggplot(aes(x=gdpPercap,y=lifeExp, color=continent, size=pop))+
  geom_jitter()+
  scale_x_log10()+
  scale_size(range = c(0.1, 10), guide = "none")+
  labs(title = "per capita GDP vs. life expectancy",
       x= "per capita GDP",
       y= "life expectancy")+
  scale_colour_brewer(palette = "Dark2")+
  facet_grid(~year)
```
8. What is the population for all countries on the Asian continent in 2007? Show this as a barplot.
```{r}
gapminder %>% 
  filter(year==2007, continent== "Asia") %>% 
  ggplot(aes(x=reorder(country, -pop), y=pop))+
  geom_bar(stat = "identity")+
  coord_flip()+
  theme(plot.title = element_text(size = rel(1.5), hjust = 0.5))
```

9. You should see that China's population is the largest with India a close second. Let's focus on China only. Make a plot that shows how population has changed over the years.
```{r}
gapminder %>% 
  filter(country=="China") %>% 
  ggplot(aes(x=year,y=pop))+
  geom_bar(stat = "identity")+
  labs(title = "Population Growth of China",
       x = "Year",
       y = "Population")
```

10. Let's compare China and India. Make a barplot that shows population growth by year using position=dodge. Apply a custom color theme using RColorBrewer.
```{r}
gapminder %>% 
  filter(country=="China"|country=="India") %>% 
  ggplot(aes(x=year,y=pop, fill=country))+
  geom_bar(stat = "identity",position="dodge")+
  labs(title = "Population Growth of China",
       x = "Year",
       y = "Population")+
   scale_fill_brewer(palette = "Dark2")
```

