---
title: "Heming HW5"
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

1. Load the data.
```{r}
library(tidyverse)
```
```{r}
life_history <- readr::read_csv("mammal_lifehistories_v2.csv")
```


2. Use your preferred function to have a look. Do you notice any problems?
```{r}
summary(life_history)
```
There are a lot of -999.

3. There are NA's. How are you going to deal with them?

I am going to replace the -999 with NA.
```{r}
life_history <- life_history%>%
  na_if("-999")
  
```


4. Where are the NA's? This is important to keep in mind as you build plots.

They are in the volumes of "mass", "gestation", "newbor", "weaning", "wean_mass", "AFR", "max_life", "litter_size", and "litters_yr".

5. Some of the variable names will be problematic. Let's rename them here as a final tidy step.
```{r}
life_history_tidy <- life_history %>%
  dplyr::rename(genus = Genus,
         wean_mass = `wean mass`,
         max_life = `max. life`,
         litter_size = `litter size`,
         litters_yr = `litters/year`) 
  
```
6. What is the relationship between newborn body mass and gestation? Make a scatterplot that shows this relationship.
```{r}
life_history_tidy %>% 
  select(newborn, gestation)
life_history_tidy
```
```{r}
life_history_tidy %>% 
  ggplot(aes(x=newborn, y=gestation))+
  geom_point()
```


7. You should notice that because of the outliers in newborn mass, we need to make some changes. We didn't talk about this in lab, but you can use scale_x_log10() as a layer to correct for this issue. This will log transform the y-axis values.


```{r}
life_history_tidy %>%
  mutate(newborn_log=log10(newborn)) %>% 
  ggplot(aes(x=newborn_log, y=gestation))+
  geom_point()
```


8. Now that you have the basic plot, color the points by taxonomic order.
```{r}
life_history_tidy %>%
  mutate(newborn_log=log10(newborn)) %>% 
  ggplot(aes(x=newborn_log, y=gestation, color=order))+
  geom_point()
```

9. Lastly, make the size of the points proportional to body mass.

```{r}
life_history_tidy %>%
  mutate(newborn_log=log10(newborn)) %>% 
  ggplot(aes(x=newborn_log, y=gestation, size=mass))+
  geom_point()
```


10. Make a plot that shows the range of lifespan by order.

```{r}
life_history_tidy %>%
  ggplot(aes(x = reorder(order, -max_life), y = max_life))+
  geom_bar(stat="identity")+
  theme(axis.text.x = element_text(angle=60, hjust=1))
```

