---
title: "Heming HW3"
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
#Load the libraries
```{r}
library(tidyverse)
```
First, load the data. Read the error message.
```{r}
fisheries <- readr::read_csv(file = "FAO_1950to2012_111914.csv")
```
1. Do you see any potential problems with the column names? Does the error message now make more sense?
Column names are duplicated so R manipulated them to unique names

2. The make.names() command is helpful when there are issues with column names. Notice that although the names are still cumbersome, much of the problemtatic syntax is removed.
```{r}
names(fisheries) = make.names(names(fisheries), unique=T) #changes the column names
names(fisheries)
```
3. Let's rename the columns. Use rename() to adjust the names as follows. Double check to make sure the rename worked correctly. Make sure to replace the old fisheries object with a new one so you can keep the column names.
country = Country..Country.
commname = Species..ASFIS.species.
sciname = Species..ASFIS.species._2
spcode = Species..ASFIS.species._1
spgroup = Species..ISSCAAP.group.
spgroupname = Species..ISSCAAP.group._1
region = Fishing.area..FAO.major.fishing.area.
unit = Measure..Measure.
```{r}
fisheries %>% 
  dplyr::rename(
    country = Country..Country.,
    commname = Species..ASFIS.species.,
    sciname = Species..ASFIS.species._2,
    spcode = Species..ASFIS.species._1,
    spgroup = Species..ISSCAAP.group.,
    spgroupname = Species..ISSCAAP.group._1,
    region = Fishing.area..FAO.major.fishing.area.,
    unit = Measure..Measure.)
```

4. Are these data tidy? Why or why not, and, how do you know?
No, we can group all the years by one colume.

5. We need to tidy the data using gather(). The code below will not run because it is commented (#) out. I have added a bit of code that will prevent you from needing to type in each year from 1950-2012 but you need to complete the remainder QQQ and remove the #.
```{r}
fisheries_tidy <- fisheries %>% 
gather(num_range('X',1950:2012), key='Year', value='Catch')
```

