---
title: "Heming 4"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Load the libraries
```{r}
library(tidyverse)
```
#gather()
Recall that we use gather() when our column names actually represent variables. A classic example would be that the column names represent observations of a variable.
```{r}
?USArrests
```
```{r}
arrests <- 
  USArrests %>% 
  mutate(State=rownames(USArrests)) %>% 
  select(State, Murder, Assault, Rape)
arrests
```
#spread()
The opposite of gather(). You use spread() when you have an observation scattered across multiple rows. In the example below, cases and population represent variable names not observations.
```{r}
country <- c("Afghanistan", "Afghanistan", "Afghanistan", "Afghanistan", "Brazil", "Brazil", "Brazil", "Brazil", "China", "China", "China", "China")
year <- c("1999", "1999", "2000", "2000", "1999", "1999", "2000", "2000", "1999", "1999", "2000", "2000")
key <- c("cases", "population", "cases", "population", "cases", "population", "cases", "population", "cases", "population", "cases", "population")
value <- c(745, 19987071, 2666, 20595360, 37737, 172006362, 80488, 174504898, 212258, 1272915272, 213766, 1280428583)

tb_data <- data.frame(country=country, year=year, key=key, value=value)
tb_data
```
When using spread() the key is the variable that you are spreading.
```{r}
tb_data %>% 
  spread(key=key, value=value)
```
#summarize()
```{r}
?msleep
```
```{r}
msleep %>% 
  filter(bodywt>=200) %>%
  summarize(mean_sleep_lg=mean(sleep_total))
```
You can also combine functions to make useful summaries for multiple variables.
```{r}
msleep %>% 
    filter(bodywt>=200) %>% 
    summarize(mean_sleep_lg = mean(sleep_total), 
              min_sleep_lg = min(sleep_total),
              max_sleep_lg = max(sleep_total),
              total = n())
```

#group_by()
```{r}
msleep %>%
  group_by(vore) %>% #we are grouping by feeding ecology
  summarize(min_bodywt=min(bodywt),
            max_bodywt=max(bodywt),
            mean_bodywt=mean(bodywt),
            total=n())
```
#Practice

1. Calculate mean brain weight by taxonomic order in the msleep data.
```{r}
msleep %>%
  group_by(order) %>% 
  summarize(mean_bodywt=mean(bodywt))
```

2. What does NA mean?
```{r}
msleep %>%
  group_by(order) %>% 
  summarize(mean_brain_wt=mean(brainwt))
```

3. Try running the code again, but this time add na.rm=TRUE. What is the problem with Cetacea?
```{r}
msleep %>%
  group_by(order) %>% 
  summarize(mean_brain_wt=mean(brainwt, na.rm=TRUE))
```
```{r}
msleep %>% 
  filter(order=="Cetacea")
```
#Load the msleep data into a new object
```{r}
msleep <- msleep
```
#Are there any NA's?
Let's first check to see if our data has any NA's. is.na() is a function that determines whether a value in a data frame is or is not an NA. This is evaluated logically as TRUE or FALSE.
```{r}
is.na(msleep)
```
OK, what are we supposed to do with that? Unless you have a small data frame, applying the is.na function universally is not helpful but we can use the code in another way. Let's incorporate it into the summarize() function.
```{r}
msleep %>% 
  summarize(number_nas= sum(is.na(msleep)))
```
This is better, but we still don't have any idea of where those NA's are in our data frame. If there were a systemic problem in the data it would be hard to determine. In order to do this, we need to apply is.na to each variable of interest.
```{r}
msleep %>% 
  summarize(number_nas= sum(is.na(conservation)))
```
What if we are working with hundreds or thousands (or more!) variables?! In order to deal with this problem efficiently we can use another package in the tidyverse called purrr.
```{r}
msleep_na <- 
  msleep %>%
  purrr::map_df(~ sum(is.na(.))) #map to a new data frame the sum results of the is.na function for all columns
msleep_na
```
Don't forget that we can use gather() to make viewing this output easier
```{r}
msleep %>% 
  purrr::map_df(~ sum(is.na(.))) %>% 
  tidyr::gather(key="variables", value="num_nas") %>% 
  arrange(desc(num_nas))
```
You can explore a specific variable more intently using count(). This operates similarly to group_by().
```{r}
msleep %>% 
  count(conservation)
```
Adding the sort=TRUE option automatically makes a descending list.
```{r}
msleep %>% 
  count(conservation, sort=TRUE)
```
#Replacing NA's
Once you have an idea of how NA's are represented in the data, you can replace them with NA so that R can better deal with them. The bit of code below is very handy, especially if the data has NA's represented as actual values that you want replaced or if you want to make sure any blanks are treated as NA.


